Computational hysteresis:\\ inter-task effects in human computation

### 2014-10-23 docs/publications/2014.09_SCIENCE/man.tex
more work on manuscript

At this point we have only discussed priming effects in terms of a 
classifier's ability to detect them.
We have said nothing qualitative about the \textit{nature} of these effects.  
This is one of the advantages of measuring priming using classifiers: one need 
not know beforehand how workers were primed, nor, 
how it might affect worker's responses.  Machine learning algorithms are
specifically designed to generalize inductively from examples, which makes
it possible to estimate how different two populations with little prior 
knowledge.

Having shown that workers are strongly primed by initial tasks, a natural
question is whether workers were biased to provide labels that reflected, 
in some way, the content of the initial tasks.  All of the experiments 
we performed involved a treatment exposed to a food-related prime, as well
as a treatment exposed to a non-food prime (objects/tools or culture), so
we can test the effects on the content of labels by determining the fraction
of food-related labels that workers provide during the test tasks.

We used wordnet, via the nltk package for python, to detect labels that are 
food-related.  Wordnet is an extensive corpus for English that contains a 
detailed set of hypernym-hyponym relationships for nouns.  Hypernyms are 
generalizations, while hyponyms are specific instances or types.  For example,
\textit{pumpernickel} is a type of \textit{bread}, so bread is a hypernym of 
\textit{pumpernickel}, and \textit{pumpernickel} is a hyponym of 
\textit{bread}.  We identified the top-level food-related entries in wordnet 
and considered all hyponym-descendants of these words to be food-related.

We use this to test the effect of initial tasks on the direction of workers
focus in both \textit{inter-t.1}, \textit{frame2}, and \textit{frame}*3.  
Recall that this experiment involved
workers either primed for the concept of food or objects/tools, using either
initial tasks or framing.  Our findings matched the straightforward
expectation that workers exposed to food related-primes are biased to 
provide more food-related labels during test tasks.  But the strength of
the bias does not account for the degree of priming we measured using a 
classifier (c.f. fig 2A).  Although we do see the expected
bias towards providing food labels, the bulk of inter-task effects would appear
to be of a more subtle nature.

Inspecting the labels manually, one striking observation was some treatments
had substantially more generic labels, for example providing labels such as
\texttt{food} or \texttt{meal} as opposed to \texttt{spaghetti}.  To test this
rigorously, we used the hypernym-hyponym relationships provided by wordnet.  
To compare the relative specificity of two sets of labels $P$ and $Q$, we first
define the relative specificity of a single label $p \in P$ in the context of
$Q$:
$$
s(p,Q) = \left\{
	\begin{array}{l l}
		\frac{N_>(p,Q) - N_<(p,Q)}{N(p,Q)} &\mathrm{if}\quad N(p,Q) > 0\\
		0 &\mathrm{else}
	\end{array}
	\right.,
$$
where $N_>(p,Q)$ is the number hyponyms of $p$ found in $Q$, and $N_<(p,Q)$ is
the number of hypernyms of $p$ found $Q$.  For this calculation, we consider 
the ``hypernyms of $p$'' to include hypernyms of hypernyms of p, etc., 
adopting the analogous interpretation for hyponyms.  Then, the relative
specificity of the entire set $P$ relative to $Q$ as:
$$
	S(P,Q) = \frac{1}{|P|}\sum_{p \in P} s(p,Q).
$$

@@

%The first two treatments, IMG:OBJ and IMG:FOOD, tested inter-task effects. 
%Workers in these treatments labelled ten images. In the IMG:OBJ treatment, the 
%\textit{initial set} of five images labeled by workers contained objects such 
%as utensils, kitchenware, and table-settings, but no food items. In the 
%\textsc{img:food}, the initial set contained centered close-ups of meals, 
%without
%any surrounding objects (except sometimes the dish supporting the food). No 
%interruption or distinction made between the initial and test sets from the 
%view of the workers. To study the positional dependance of inter-task effects,
%these treatments were subdivided, and workers were shown one of five 
%permutations. There are of course 5! = 120 permutations of five images, but 
%investigating all permutations was too expensive, and probably redundant, so 
%we chose 5 permutations that ensure each image occupies each position within 
%the test set. We discuss the benefits and drawbacks of this choice in the 
%supplementary material.
%
%Workers in the treatments \textsc{frm:food} and \textsc{frm:obj} were shown a 
%framing slide, which bore a message reading: ``Funded by the laboratory for 
%the visual perception of \{Food and Ingredients $\vert$ Objects and Tools\}''. 
%After the framing slide, workers in these treatments proceeded directly to 
%the test set (and so labeled only 5 images).

\paragraph{Inter-task effects and specificity.}
Still using the results of the first experiment, we investigated how initial 
tasks affect the degree of specificity during the test tasks.
To measure the relative specificity of labels, we used the hypernym-hyponym
relationships provided by the wordnet corpus.  Hypernym relationships point
from specific types to generalizations, while hyponyms point in the opposite
direction.  For example, the word \texttt{bread} has the hypernym 
\texttt{baked-goods}, and the hyponym \texttt{naan}.
We calculated the relative specificity $s_i(u,v)$ of the labels attributed by 
two workers $u$ and $v$ to image $i$ as follows:
$$
	s_i(u,v) = \sum_{\ell \in u(i)} \sum_{\ell \in v(i)} 
	\left( \mathbf{1}_{[\ell > m]} - \mathbf{1}_{[m>\ell]} \right),
$$
where $u(i)$ denotes the set of labels attributed by worker $u$ to image $i$, 
and $\mathbf{1}_{[l>m]}$ evaluates to 1 if $\ell$ is more specific than $m$
(meaning that $m$ can be reached from $\ell$ by following hypernyms), 
and 0 otherwise. We then define the relative specificity $S_i(U,V)$ of two 
treatments $U$ and $V$ relative to image $i$ to be the mean 
relative specificity of two uniformly drawn workers:
$$
S_i(U,V) = \mathrm{E}_{u,v}\{s_i(u,v)\}   
$$
$$
\hat{S}_i(U,V) = \frac{1}{|U \times V|} \sum_{u \in U} \sum_{v in V}
		s_i(u,v).
$$
Looking at labels attributed to the first test image, we found workers 
that initially labeled images of ingredients were more specific 
than those that initially labeled cultural scenes (see \textbf{Fig. 3}).  One 
possible reason for
this is that the images of ingredients directed workers' attention toward
the component parts in the images.  To try to corroborate this, we can look
at the specificity of workers whose initial tasks contained images of prepared
meals.  Since these images contained whole prepared meals, rather than 
separated ingredients, we can isolate the effects due to showing the presence
of food in the initial images from the presentation of isolated parts.

In fact, the workers that intially labeled prepared meals were even
more specific than those that initially labeled ingredients.  Recall that the
test tasks also depicted prepared meals.  Thus, it would appear that it is 
the similarity of initial and test tasks which drove similarity, and not the
depiction of isolated parts (ingredients).

@@

Such a phenomenon would be consistent with the psychological mechanism known as
\textit{negative priming}.
Negative priming occurs when a person becomes desensitized to non-salient 
stimuli to which she is repeatedly exposed. Consider that, workers that 
initially labeled images depicting prepared meals might no longer
regard the generic labels \texttt{food} or \texttt{meal} to be salient, and 
opt instead for \texttt{bread}, or \texttt{pasta}.  But for workers that
initially labeled cultural images, these rather generic labels might well seem
appropriate.

\paragraph{The role of task position in inter-task effects.}
Up to this point, we have focused on the inter-task effects induced on 
the first test task.  Not suprisingly, this task was, in general, more 
strongly affected than the others.  It would stand to reason that, as workers 
proceed through the test images, priming induced by the initial tasks would
be ``washed out'', and replaced by the effects of the test tasks themselves,
which do not differ between treatments.

>>>>

At this point we have only discussed priming effects in terms of a 
classifier's ability to detect them.
We have said nothing qualitative about the \textit{nature} of these effects.  
This demonstrates an advantage of using classifiers to measure priming.  
Machine learning algorithms are designed to find salient differences 
by generalizing, inductively, from examples, so one
need not know beforehand how workers were primed, nor the natuer of the  
priming effects.  Machine learning algorithms are specifically designed to 
generalize inductively from examples, which makes
it possible to estimate how different two populations with little prior 
knowledge.

But once having shown that workers are strongly primed by initial tasks, it is
natural ask whether workers were biased to provide labels that reflected, 
in some way, the \textit{content} of the initial tasks.  The conceptual basis
for the primes in experiment 2 were eithre \textit{food} or 
\textit{objects and tools}.  Thus, we can test whether workers exposed to food
primes produced more food-related.

To identify food-related labels, we used the wordnet corpus.  Wordnet
is an extensive corpus for English, that includes hypernym-hyponym 
relationships between words.  The hypernym relation correpsonds to conceptual
abstraction.  For example \textit{bread} is a hypernym of 
\textit{pumpernickel} because pumpernickel is a \textit{type of} bread.
The hyponym relation is simply the converse: pumpernickel is a hyponym
of bread.  For the purpose of our analysis, when we refer to hypernyms,
we will mean both direct hypernyms, as well as the hypernyms of hypernims,
and so on, with an analogous interpretation for hyponyms.

We collectod a small set of top-level food-related nouns in wordnet,
together with all of their hyponyms, and took the resulting set of 
3589 lemmas as an operational definition for the set of English food-related 
words.  This set is quite extensive, containing somewhat obscure forms such as
``burgoo'' and ``barmbrack'', and consumables at the fringe of what might
be considered food such as ``vitamin B'' (or if you prefer, 
``B complex''), as well as more common lemmas like
``banana'',  ``bread'', and of course ``banana bread''.  We found that,
for the data in experiment 1, this method produced excellent aggreement with 
human annotation of food-related words (see supplementary material).  
For the analysis exp2, the variety of non-English names for ethnic dishes 
forced us to rely on a manual approach.  This corroborated our results
for exp1, and we encourage the interested reader to read our discussion of
the manual approach in the supplementary material.

Our findings matched the straightforward expectation that workers exposed to 
food related-primes were biased to provide more food-related labels during 
test tasks (see Fig. 3A).  Had we relied on the intuition that workers 
should be biased to label according to the underlying concept of their prime,
we would have underestimated the strength of priming effects.  This
again illustrates the advantage of using classifiers to detect priming as
opposed to testing for the effects that one might intuitively expect.  
Though present, the expected influence on the number of food-related words
cannot account for the bulk of inter-taks effects.

\paragraph{Effects on label vocabulary.}
Inspecting the labels manually, one striking observation was that in
\textit{test-t.2}, workers that initially images of cultural scenes seemed
more likely to give generic labels, compared to those that initially labeled 
images depicting prepared meals.  Recall that, for these treatments, the test 
tasks contained images depicting prepared meals, which are very similar in
content to the initial tasks for the latter set of workers.  Workers that
had already labeled 5 images depicting food were less likely to attribute
the generic label ``food'' during the first test task, doing so only 16 times,
compared to 76 times by workers that had labeled culture-related images 
containing no food.  This observation suggests that, when workers label a 
series of very similar images.

One straightforward way to test whether workers from exp2.task.food are 
being more specific than those from exp2.taks.cult is to look at the number 
of unique labels they provided for each image.  Workers attributing more
specific labels should produce a larger vocabulary of unique labels for a 
given, which is indeed what we found (see Fig. 3B).  Workers from 
exp2.task.food produced as much as 33\% more unique labels per image than 
workers from exp2.task.cult, and the relative excess of the vocabulary size
followed the general trend of inter-task effects that we observed in 
Fig. 2B.

To confirm our expectation that the larger vocabulary corresponds to a 
the use of \textit{more nuanced} words, we used the 
hypernym-hyponym of wordnet to operationalize a notion of relative 
specificity.  We define the relative specificity of two sets of labels $P$ 
and $Q$ as:
$$
	S(P,Q) = \frac{1}{|P \times Q|}\sum_{p \in P} \sum_{q \in Q} 
		\left(\mathbf{1}_{[p>q]} - \mathbf{1}_{[q>p]}\right),
$$
where $\mathbf{1}_{[p>q]}$ evaluates to 1 if $p$ is a hyponym of $q$.
Using the above equation, we found that, for every images in the 
test set, the labels attributed by workers in exp2.task.food were more 
specific than those by workers in exp2.task.cult.

Thus, when workers perform a series of similar image labeling tasks it would
appear that they tend to avoid generic words, opting instead for more specific
ones, and yielding a richer vocabulary of labels.  Qualitative 
characterisation and classification is one of the most comman applications of 
microtask platforms, and this finding suggests that bundling similar tasks 
together substantially increases the specificity of HPU output.

Our observations of the effects on specificity would appear to be consistent
with the phenomenon of negative priming.
Negative priming occurs when a person becomes desensitized to non-salient 
stimuli to which she is repeatedly exposed. Consider that, workers that 
initially labeled images depicting prepared meals might no longer
regard the generic labels \texttt{food} or \texttt{meal} to be salient, and 
opt instead for \texttt{bread}, or \texttt{pasta}.  But for workers that
initially labeled cultural images, these rather generic labels might well seem
appropriate.

@@

@@

Such a phenomenon would be consistent with the psychological mechanism known as
\textit{negative priming}.
\paragraph{The role of task position in inter-task effects.}
Up to this point, we have focused on the inter-task effects induced on 
the first test task.  Not suprisingly, this task was, in general, more 
strongly affected than the others.  It would stand to reason that, as workers 
proceed through the test images, priming induced by the initial tasks would
be ``washed out'', and replaced by the effects of the test tasks themselves,
which do not differ between treatments.
